Rheology of Cubic Blue Phases

### 2012-11-08 trunk/papers/bprheo/bprheo.tex
Updated text - version for Davide to check out
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@1809 40876441-3496-0410-a613-a2fca35b2ae2

It is possible to define a dimensionless number that describes the deformation
of the director field under flow. The so-called Ericksen number
is defined as 
\begin{equation}
Er=\frac{\eta V L}{K}
\end{equation}
with $\eta$ and $K$ being the viscosity and elastic constant and
$V$ and $L$ as typical velocity and length scale.

@@

\begin{table*}[htpb]
\begin{tabular}{|c||c|| c || c || c |c |c||c| c| c||c| c| c|}
\hline
& $\dot{\gamma}$ & ${\it Er}$ & $q_0$ & $\bar{v}_{x,min}$ & $\bar{v}_{x,max}$ & $\bar{v}_{x,std}$ & $\bar{v}_{y,min}$ & $\bar{v}_{y,max}$ & $\bar{v}_{y,std}$ & $\bar{v}_{z,min}$ & $\bar{v}_{z,max}$ & $\bar{v}_{z,std}$ \\
\hline
BPI \\
\hline
AN & 0.244 & 0.08& 0.1388 &-15.79 &15.57 &1.04 &-0.89 &0.91 &0.95 &-1.59 &1.19 &1.27 \\
AN & 0.488 & 0.17 & 0.1388 &-31.11 &31.25 &1.31 &-0.93 &0.98 &1.28 &-1.62 &1.10 &1.40 \\
AN & 0.976 & 0.33 & 0.1388 &-62.24 &62.37 &2.51 &-1.31 &1.25 &2.68 &-1.24 &0.87 &2.65 \\
AN & 1.95 & 0.67 & 0.1388 &-124.65 &124.82 &3.62&  -2.49 &2.50 &4.51 &-1.89 & 1.62 &3.51 \\
\hline
3D-PRC & 3.91 & 1.33 & 0.1388 &-248.69 &248.74 &3.90&  -3.31 &3.35 &4.21 &-2.56 & 2.88 &4.39 \\
3D-PRC & 7.81 & 2.67 & 0.1388 &-497.22 &496.16 &4.46 &-6.49 &6.56 &7.99 &-5.31 & 7.46 &6.81 \\ 
3D-PRC & 15.63 & 5.33 & 0.1388 &-992.46 &993.18 &10.36 &-3.08 &3.17 &10.49 &\bf{-2.87} & \bf{3.57} &10.54 \\
3D-PRC & 15.63 & 5.33 & -0.1388 &-992.46 &993.18 &10.36 &-3.08 &3.17 &10.49 &\bf{-3.57} & \bf{2.87} &10.54 \\
3D-PRC & 31.25 &10.67 & 0.1388 & -1988.36 &1986.71 &16.85 &-4.08 &4.46 &15.94 &-11.37 & 12.16 &19.38\\
3D-PRC & 46.87 & 15.99 & 0.1388 & -3025.46 & 3003.05 &21.43 & -56.96 & 56.44  & 18.75 & -12.65 & 82.97  & 27.96\\
\hline
2D-DTR & 62.50 & 21.33& 0.1388 & -4039.41 &3995.3  & 20.57 & -62.63 & 62.13 & 24.68 &-73.52 & 110.76 & 33.26 \\
2D-DTR &78.125 &26.66 & 0.1388 & -4968.11 &4973.75 &20.89 &-60.61 &58.75 &23.66 &-84.93 &116.61 &34.75 \\
\hline
1D-CH & 93.75 &32.00 & 0.1388 & -5955.12 &5948.84 & 0.71 & -0.002 & 0.002 & -  & -16.45 & 16.03 &0.62 \\
1D-CH & 125.0 & 42.66& 0.1388 & -7934.97 & 7935.41 & 0.36 & -0.005 & 0.005 & - &-22.73 & 21.12 & 0.07\\
\hline
FAN & 187.5 & 63.99& 0.1388 &-11906.4  &119064 &0.36 &- &- &- &-0.68 & 0.38 & 0.05\\
\hline

BPII \\
\hline
3D-PRC & 0.244 &0.04 & 0.1963 &-7.75 &7.70 &0.13 &-0.17 &0.17 &0.13 &-0.24 &0.23 &0.19 \\
3D-PRC & 0.488 &0.08 & 0.1963 &-15.50 &15.40 &0.19 &-0.32 &0.30 &0.21 &-0.43 &0.42 &0.29 \\
3D-PRC & 0.976 &0.17 & 0.1963 &-31.01 &30.80 &0.33 &-0.60 &0.58 &0.41 &-0.85 &0.79 &0.47 \\
3D-PRC & 1.95 &0.33  & 0.1963 &-62.04  &61.60 & 0.97 & -1.09 &1.07 & 0.76 & -1.64 & 1.55 & 0.81\\
3D-PRC &3.91 & 0.67& 0.1963 &-123.93 &123.18 & 1.76 &-1.66 &1.70 & 1.45 &-3.09& 2.73 &1.47\\
3D-PRC &7.81 &1.33 & 0.1963  &-247.67 &246.45 & 3.20 &-2.62 &2.71 & 2.67 &-5.78 & 4.77 &2.74\\
3D-PRC &15.63 & 2.67& 0.1963 &-494.85 &493.09 & 5.64 &-3.13 &3.31 &4.24 &\bf{-10.00} & \bf{7.66} &4.33\\
3D-PRC & 15.63 & 2.67& -0.1963&-494.85 &493.09 & 5.64 & -3.13 &3.31 &4.24 &\bf{-7.66} & \bf{10.00} &4.33\\
3D-PRC & 31.25 &5.34 & 0.1963 &-988.73 &985.93 &8.96  &-1.79 &1.82 &5.86 &-14.39 & 11.04 &6.35\\
\hline
1D-THW & 46.87 &8.00 & 0.1388 & -1476.75 & 1476.84 &1.69 &-0.35  &0.35  & 0.17 &-0.23 &0.23  & 0.22\\
1D-THW & 62.50 & 10.67& 0.1963 & -1969.41  & 1969.41 & 3.07 & -1.20 & 1.20 & 0.38 &-0.96 & 1.48 &0.38 \\
\hline
1D-CH & 78.125 & 13.33 &  0.1388 &-2458.60 & 2458.61  &0.37 &- &- &- &-18.54 &16.94 &0.32 \\
1D-CH & 93.75 &16.00 & 0.1388 &-2950.2  &2950.2 &0.50 &- &- &- &-22.69 & 19.72 & 0.04\\
\hline
FAN & 125.0 &21.33 & 0.1388 &-3937.31  & 3937.32 &0.28 &-  &-  &-  &-1.79 &1.15  &0.14 \\
\hline
\end{tabular}
\caption{Minima, maxima and standard deviation of time-averaged velocity components 
for BPI ($\tau=-0.5, \kappa=1.0$) and BPII ($\tau=-0.5, \kappa=2.0$) in steady 
shear flow: All values are given in $10^{-5}$ lattice units. The transient dynamics 
has not been included in the averaging procedure. The regimes comprise periodically 
recurring conformation (3D-PRC), two-dimensional double twist rolls (2D-DTR), 
one-dimensional helix (1D-CH), one-dimensional travelling helical wave (1D-THW) 
and a flow-aligned nematic state (FAN).}
\label{tab1}
\end{table*}

@@

For shear rates $\gd\ge6.25\e{-4}$ the network of disclination lines
breaks up and oscillations in the stress signal are absent.
In this regime of shear rates, to which we refer as BPII-2, the 
blue phase dissolves into a simple cholesteric liquid 
crystal. The mode of flow is a travelling helical wave with the helical axis 
oriented along the vorticity direction.

@@

If the flow rate exceeds a certain critical value BPII does not undergo the 
affine transformation  above and adopts another mode of flow. 
Fig. \ref{bp2-high} shows a sequence of cuts through the director 
field at high shear rate. The entire blue phase network has broken up 
into a simple cholesteric helix and flows as a travelling 
helical wave with the helical axis oriented along the vorticity direction.
The state is translationally invariant in flow and gradient direction.
and is the preferred mode of flow of a cholesteric 
liquid crystal at low Ericksen numbers below the uncoiling transition 
to a nematic, flow-aligned state at even higher shear rates ~\cite{Rey:1996a, Rey:1996b}.

In our scenario the break-up occurs for shear rates $\gd>3.12\e{-4}$, which
corresponds to Ericksen numbers of ${\it Er}\simeq10$, based on our
parameters ($K=0.02, \eta=0.66$), typical velocities (viz. Tab. \ref{tab1}) 
and length scales ($p/2$).

@@

@@

BPI has body-centered symmetry and unlike BPII. The disclination lines
 are well separated and do not intersect.
We believe that this topological difference is responsible for most of
the differences between BPI and BPII regarding their rheological response.

We present again key aspects in an overview before we 
address some features in more detail. Fig. \ref{bp1-rheo} shows the 
apparent viscosity $\eta_{app}$ versus time and strain for the same shear rates
as in Fig. \ref{bp2-rheo}.
For shear rates in regime BPI-3, i.e. $\gd\ge6.25\e{-4}$ a breakup of 
the disclination network occurs. This is the same critical shear rate as 
the one observed in BPII-2. But this time the configuration is quasi 
two-dimensional.

In regime BPI-2 at lower shear rates periodic oscillations emerge in the 
deviatoric stress that are not sinusoidal, but show an increasingly complex
pattern up to the breakup point.
For even lower shear rates $\gd\le3.91\e{-5}$ in regime BPI-1 
these oscillations become very erratic. There is no regular recurrence and 
after a few cycles the stress pattern is completely irregular. 
Shear-thinning is slightly stronger than in BPII. 
The average shear stress during one cycle can be fitted reasonably 
well by a power law $\bar{\Pi}_{xy}=a \gd^b$ with $a=0.019, b=0.63$.

@@

\begin{figure*}[htpb]
\includegraphics[width=0.245\textwidth]{disc-xy-10k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-180k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-200k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-210k_run1115.png}\\
\includegraphics[width=0.245\textwidth]{disc-xy-400k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-700k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-750k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-2200k_run1115.png}
\caption{Snapshots of BPI disclination network at $\dot{\gamma}=4.88\e{-6}$: 
Depicted is the transition from flow-induced, intertwined helices that 
undergo a recurring affine transformation (top left and centre, both at 
time step $t=4.0\e{5}$; top right at time step $t=7.0\e{5}$) with an intermediate 
conformation that is non-helical (bottom left at time step $t=7.5\e{5}$) 
to a distorted state (bottom centre at time step $t=1.7\e{6}$) and 
finally to an amorphous network (bottom right at $t=2.2\e{6}$). 
The picture at the top centre shows the network in gradient direction.
All other pictures have the view along the vorticity direction.}
\label{bp1-low}
\end{figure*}

@@

Adjacent to regime BPI-1 ($\gd<3.91\e{-5}$) and 
at slightly larger shear rates $3.91\e{-5}\le\gd< 6.25\e{-4}$
and Ericksen numbers $0.5\le{\it Er}<10$ 
lies another region where the network flows with periodically 
recurring conformations (Fig. \ref{bp1-rheo}). 
We refer to this region as BPI-2. The transition between BPI-1 and BPI-2 
is shown in Fig. \ref{bp1-1_bp1-2} by means of the deviatoric stress versus total strain. 
A qualitative difference between these two is the absence of 
large stress fluctuations that occur during one typical cycle in regime BPI-1.
Remember that for even lower $\gd$ the deviatoric stress becomes 
temporarily negative (Fig. \ref{bp1-fe-yield}), which 
caused destabilisation of the periodic network and led to the 
amorphous configuration in the steady state. Hence, an explanation 
for this phenomenon could be that in BPI-2 the network finds a 
kinetic path to bypass these large amplitude fluctuations that 
prevents the network from destabilising.

@@

\begin{figure}[htpb]
\includegraphics[width=0.45\textwidth]{dir3d-z-302k_run916.png}
\caption{Director field of BPI at high shear rate: The pictures show a slice 
of $L_x\times L_y \times L_Z= 32\times32\times 32$ lattice sites along the vorticity direction 
with the flow/gradient direction oriented along the horizontal/vertical axis. 
The colour code gives the magnitude of the in-plane component of the director field. 
The data is translationally invariant in vorticity direction.}
\label{bp1-high}
\end{figure}

At high flow rates we observe a break-up of the flow-induced 
disclination network of regime BPI-2. However, 
contrary to BPII, where the configuration is a travelling helical 
wave (viz. Fig. \ref{bp2-high}), BPI prefers a quasi 
two-dimensional, translationally invariant state, which is shown in 
Fig. \ref{bp1-high}. The snapshots visualise the affine 
transformation of simple shear flow in the gradient-flow plane.
There are distinctive regions where the director field is 
predominantly oriented along the vorticity direction and 
other regions where it is oriented rather in flow-gradient plane. 
Occasionally while moving past each other these rolls resemble 
regions of local double twist.
Interestingly the critical shear rate for the break-up lies in the same 
window as that of transition between BPII-1 and BPII-2, 
 $3.125\e{-4}\le\gd_c\le 6.25\e{-4}$ and occurs at similar 
Ericksen numbers ${\it Er}\simeq 10$.

\subsection{Transition to cholesteric helix and flow-aligned nematic state}

\begin{figure}[htpb]
\includegraphics[width=0.45\textwidth]{dir3d+y-200k_run1179.png}
\caption{}
\label{cholflow}
\end{figure}

\clearpage

@@

We were able to characterise the rheology of cubic BPs and identified 
two different flow regimes for BPII and three different ones in case of BPI. 
Below Ericksen numbers ${\cal O}(10)$ BPII exhibits weak shear-thinning and 
obeys a power-law.  The BPII disclination network breaks up and reconnects 
in the flow, which leads to a periodically recurring dependence 
of the deviatoric shear stress. The flow-induced conformation look generally
very similar to the quiescent network at equilibrium. 
While being affinely transformed due to the shear flow
the disclination network moves in the vorticity direction. The sense of
motion is directly linked to the helicity of the underlying cholesteric phase.

At larger Ericksen numbers the flowing BPII network breaks up completely into 
a simple cholesteric phase and flows as a travelling helical wave.

Interestingly, BPI shows a flow behaviour that is very different to that of BPII.
This is a direct consequence of the topological differences in both disclination 
network. Below Ericksen numbers ${\cal O}(0.1)$ the BPI network cannot flow and 
retain a regular appearance. Shortly after the onset of the shearing 
very large stress fluctuations occur. At the lowest shear rates $\gd<9.76\e{-6}$
the contribution of the network to the total deviatoric stress is actually temporarily 
negative during one cycle. These fluctuations destabilise the network and 
eventually trigger its dissolution into an amorphous state with 
dynamic yield stress.

At slightly larger Ericksen numbers $0.1<{\it Er}<10$ shear thinning is 
slightly stronger than for BPII, but BPI does not show simple power-law 
behaviour. 
Instead, it appears BPI kinematically bypasses the large stress fluctuations and 
flows with very complex, periodically recurring flow-induced conformations.
The offset of the network after each cycle is random in magnitude and direction.
However, despite their complex appearance the flow-induced conformations are 
topologically connected to the quiescent BPI and switching off of the 
shear flow often restores a perfect equilibrium.
It is possible to interpret these findings as an instance of 
deterministic rheochaos.

Just as in case of BPII, the BPI network breaks up at Ericksen numbers ${\cal O}(10)$,
but adopts a quasi-two-dimensional configuration that is translationally 
invariant in vorticity direction.

>>>>

It is possible to define a dimensionless number that describes the deformation
of the director field under flow. The so-called Ericksen number
is defined as 
\begin{equation}
Er=\frac{\eta V L}{K}
\end{equation}
with $\eta$ and $K$ being the viscosity and elastic constant and
$V$ and $L$ as typical velocity and length scale. 
In the present work $L=p_0/2$ was assumed to be the unit cell size,
whereas $V$ was taken to be the velocity difference
across one unit cell, i.e. $V=\dot{\gamma} p_0/2$.

@@

\begin{table*}[htpb]
\begin{tabular}{|c||c|| c || c || c |c |c||c| c| c||c| c| c|}
\hline
& $\dot{\gamma}$ & ${\it Er}$ & $q_0$ & $\bar{v}_{x,min}$ & $\bar{v}_{x,max}$ & $\bar{v}_{x,std}$ & $\bar{v}_{y,min}$ & $\bar{v}_{y,max}$ & $\bar{v}_{y,std}$ & $\bar{v}_{z,min}$ & $\bar{v}_{z,max}$ & $\bar{v}_{z,std}$ \\
\hline
BPI \\
\hline
BPI-1 & 0.244 & 0.04& 0.1388 &-15.79 &15.57 &1.04 &-0.89 &0.91 &0.95 &-1.59 &1.19 &1.27 \\
BPI-1 & 0.488 & 0.08 & 0.1388 &-31.11 &31.25 &1.31 &-0.93 &0.98 &1.28 &-1.62 &1.10 &1.40 \\
BPI-1 & 0.976 & 0.17 & 0.1388 &-62.24 &62.37 &2.51 &-1.31 &1.25 &2.68 &-1.24 &0.87 &2.65 \\
BPI-1 & 1.95 & 0.33 & 0.1388 &-124.65 &124.82 &3.62&  -2.49 &2.50 &4.51 &-1.89 & 1.62 &3.51 \\
\hline
BPI-2 & 3.91 & 0.67 & 0.1388 &-248.69 &248.74 &3.90&  -3.31 &3.35 &4.21 &-2.56 & 2.88 &4.39 \\
BPI-2 & 7.81 & 1.33 & 0.1388 &-497.22 &496.16 &4.46 &-6.49 &6.56 &7.99 &-5.31 & 7.46 &6.81 \\ 
BPI-2 & 15.63 & 2.67 & 0.1388 &-992.46 &993.18 &10.36 &-3.08 &3.17 &10.49 &\bf{-2.87} & \bf{3.57} &10.54 \\
BPI-2 & 15.63 & 2.67 & -0.1388 &-992.46 &993.18 &10.36 &-3.08 &3.17 &10.49 &\bf{-3.57} & \bf{2.87} &10.54 \\
BPI-2 & 31.25 &5.33 & 0.1388 & -1988.36 &1986.71 &16.85 &-4.08 &4.46 &15.94 &-11.37 & 12.16 &19.38\\
BPI-2 & 46.87 & 8.00 & 0.1388 & -3025.46 & 3003.05 &21.43 & -56.96 & 56.44  & 18.75 & -12.65 & 82.97  & 27.96\\
\hline
BPI-3 & 62.50 & 10.67& 0.1388 & -4039.41 &3995.3  & 20.57 & -62.63 & 62.13 & 24.68 &-73.52 & 110.76 & 33.26 \\
BPI-3 &78.13 &13.33 & 0.1388 & -4968.11 &4973.75 &20.89 &-60.61 &58.75 &23.66 &-84.93 &116.61 &34.75 \\
\hline
CHG & 93.75 &16.00 & 0.1388 & -5955.12 &5948.84 & 0.71 & -0.002 & 0.002 & -  & -16.45 & 16.03 &0.62 \\
CHG & 125.0 & 21.33& 0.1388 & -7934.97 & 7935.41 & 0.36 & -0.005 & 0.005 & - &-22.73 & 21.12 & 0.07\\
\hline
FAN & 187.5 & 32.00& 0.1388 &-11906.4  &119064 &0.36 &- &- &- &-0.68 & 0.38 & 0.05\\
\hline

BPII \\
\hline
BPII-1 & 0.244 &0.02 & 0.1963 &-7.75 &7.70 &0.13 &-0.17 &0.17 &0.13 &-0.24 &0.23 &0.19 \\
BPII-1 & 0.488 &0.04 & 0.1963 &-15.50 &15.40 &0.19 &-0.32 &0.30 &0.21 &-0.43 &0.42 &0.29 \\
BPII-1 & 0.976 &0.08 & 0.1963 &-31.01 &30.80 &0.33 &-0.60 &0.58 &0.41 &-0.85 &0.79 &0.47 \\
BPII-1 & 1.95 &0.17  & 0.1963 &-62.04  &61.60 & 0.97 & -1.09 &1.07 & 0.76 & -1.64 & 1.55 & 0.81\\
BPII-1 &3.91 & 0.33& 0.1963 &-123.93 &123.18 & 1.76 &-1.66 &1.70 & 1.45 &-3.09& 2.73 &1.47\\
BPII-1 &7.81 &0.67 & 0.1963  &-247.67 &246.45 & 3.20 &-2.62 &2.71 & 2.67 &-5.78 & 4.77 &2.74\\
BPII-1 &15.63 & 1.33& 0.1963 &-494.85 &493.09 & 5.64 &-3.13 &3.31 &4.24 &\bf{-10.00} & \bf{7.66} &4.33\\
BPII-1 & 15.63 & 1.33& -0.1963&-494.85 &493.09 & 5.64 & -3.13 &3.31 &4.24 &\bf{-7.66} & \bf{10.00} &4.33\\
BPII-1 & 31.25 &2.67 & 0.1963 &-988.73 &985.93 &8.96  &-1.79 &1.82 &5.86 &-14.39 & 11.04 &6.35\\
\hline
BPII-2 & 46.87 &4.00 & 0.1388 & -1476.75 & 1476.84 &1.69 &-0.35  &0.35  & 0.17 &-0.23 &0.23  & 0.22\\
BPII-2 & 62.50 & 5.34& 0.1963 & -1969.41  & 1969.41 & 3.07 & -1.20 & 1.20 & 0.38 &-0.96 & 1.48 &0.38 \\
\hline
CHG & 78.13 & 6.66 &  0.1388 &-2458.60 & 2458.61  &0.37 &- &- &- &-18.54 &16.94 &0.32 \\
CHG & 93.75 &8.00 & 0.1388 &-2950.2  &2950.2 &0.50 &- &- &- &-22.69 & 19.72 & 0.04\\
\hline
FAN & 125.0 &10.67 & 0.1388 &-3937.31  & 3937.32 &0.28 &-  &-  &-  &-1.79 &1.15  &0.14 \\
\hline
\end{tabular}
\caption{Minima, maxima and standard deviation of time-averaged velocity components 
for BPI ($\tau=-0.5, \kappa=1.0$) and BPII ($\tau=-0.5, \kappa=2.0$) in steady 
shear flow: All values are given in $10^{-5}$ lattice units. The transient dynamics 
has not been included in the averaging procedure. The regimes comprise the formation
of an amorphous network (BPI-1), periodically recurring conformations (BPI-2 and BPII-1), 
two-dimensional double twist rolls (BPI-3), a one-dimensional travelling helical wave (BPII-2), 
a cholesteric helix oriented along the gradient direction (CHG), 
and a flow-aligned nematic state (FAN).}
\label{tab1}
\end{table*}

@@

For shear rates $4.68\e{-4}\lesssim\gd\lesssim6.25\e{-4}$ the network of disclination lines
breaks up and oscillations in the stress signal are absent.
In this regime of shear rates, to which we refer as BPII-2, the 
blue phase dissolves into a simple cholesteric liquid 
crystal. The helical axis is oriented in vorticity direction and 
the mode of flow is a travelling helical wave.

At shear rates $7.81\e{-4}\lesssim\gd9.37\e{-4}$ the helix changes orientation.
The helical axis is now oriented along the gradient direction with
the liquid crystal flowing in 'nematic planes'. For even higher shear rates
the system undergoes a transition to a flow-aligned nematic state.

@@

If the flow rate exceeds a certain critical value BPII does not undergo the 
affine transformation  above and adopts another mode of flow. 
Fig. \ref{bp2-high} shows a sequence of cuts through the director 
field at high shear rate. The entire blue phase network has broken up 
into a simple cholesteric helix and flows as a travelling 
helical wave with the helical axis oriented along the vorticity direction.
The state is translationally invariant in flow and gradient direction.
and is the preferred mode of flow of a cholesteric 
liquid crystal in vorticity mode at low Ericksen numbers, i.e. below 
the uncoiling transition to a nematic, flow-aligned state at even 
higher shear rates ~\cite{Rey:1996a, Rey:1996b}.

In our scenario the break-up occurs for shear rates $\gd\simeq 4.68\e{-4}$, which
corresponds to Ericksen numbers of ${\it Er}\simeq4$, based on our
parameters for BPII ($K=0.02, \eta=0.666$, Tab. \ref{tab1} and $p_0/2=16$).

@@

We observe another configuration between the travelling helical wave and the 
flow-aligned nematic state at shear rates $7.81\e{-4}\lesssim\gd\lesssim9.37\e{-4}$, 
which has so far not been reported to our knowledge \com{CHECK}.
We will address this in section \ref{cholflow}.

@@

BPI has body-centered symmetry and unlike BPII. The disclination lines
 are well separated and do not intersect.
We believe that this topological difference is responsible for most of
the differences between BPI and BPII regarding their rheological response. 
We present again key aspects in an overview before we 
address some features in more detail.

Fig. \ref{bp1-rheo} shows the apparent viscosity $\eta_{app}$ versus time and 
strain for the same shear rates as in Fig. \ref{bp2-rheo}.
For shear rates $9.37\e{-4}\lesssim\gd\lesssim1.25\e{-3}$ we observe the same cholesteric 
configuration with the helical axis along the gradient direction as for BPII (see
section \ref{cholflow}).

In regime BPI-3 $6.25\e{-4}\lesssim\gd\lesssim7.81\e{-4}$ a breakup of 
the blue phase disclination network occurs. This is about the same critical shear rate 
where we observe the transition to BPII-2. However, for BPI the configuration is not
a travelling helical wave, but rather a quasi two-dimensional state consisting of
double-twist rolls.

At shear rates $3.91\e{-5}\lesssim\gd\lesssim4.68\e{-4}$ we find flow with 
periodically recurring conformations. We refer to this regime as BPI-2. 
The oscillations in the 
deviatoric stress that are not sinusoidal, but show an increasingly complex
pattern up to the breakup point.
For even lower shear rates $\gd\lesssim3.91\e{-5}$ in regime BPI-1 
these oscillations become very erratic and unstable. 
There is no regular recurrence and 
after a few cycles the stress pattern is completely irregular.
Finally, the system ends up in an amorphous configuration.
Shear-thinning is slightly stronger than in BPII.
The average shear stress during one cycle can be fitted reasonably 
well by a power law $\bar{\Pi}_{xy}=a \gd^b$ with $a=0.019, b=0.63$.

@@

\begin{figure*}[htpb]
\includegraphics[width=0.245\textwidth]{disc-xy-10k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-180k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-200k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-210k_run1115.png}\\
\includegraphics[width=0.245\textwidth]{disc-xy-400k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-700k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-750k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-2200k_run1115.png}
\caption{Snapshots of BPI disclination network at $\dot{\gamma}=4.88\e{-6}$: 
Depicted is the transition from the quiescent state to flow-induced, intertwined 
helices that undergo a recurring affine transformation. The top row
shows a section of one unit cell for early times 
$t=1\e{4}, 1.8\e{5}, 2.0\e{5}$ and $2.1\e{5}$. The bottom 
pictures on the far left, centre left and centre right 
show the situation at later time steps $t=4.0, 7.0$ and $7.5\e{5}$,
which the system passes for several cycles before it ends up in
an amorphous state (bottom row far right at time step $t=2.2\e{6}$).}
\label{bp1-low}
\end{figure*}

@@

Adjacent to regime BPI-1 ($\gd<3.91\e{-5}$) and 
at slightly larger shear rates $3.91\e{-5}\le\gd< 6.25\e{-4}$
and Ericksen numbers $0.67\le{\it Er}<8$ 
lies another region where the network flows with periodically 
recurring conformations (Fig. \ref{bp1-rheo}). 
We refer to this region as BPI-2. The transition between BPI-1 and BPI-2 
is shown in Fig. \ref{bp1-1_bp1-2} by means of the deviatoric stress versus total strain. 
A qualitative difference between these two is the absence of 
large stress fluctuations that occur during one typical cycle in regime BPI-1.
Remember that for even lower $\gd$ the deviatoric stress becomes 
temporarily negative (Fig. \ref{bp1-fe-yield}), which 
caused destabilisation of the periodic network and led to the 
amorphous configuration in the steady state. Hence, an explanation 
for this phenomenon could be that in BPI-2 the network finds a 
kinetic path to bypass these large amplitude fluctuations that 
prevents the network from destabilising.

@@

\begin{figure}[htpb]
\includegraphics[width=0.45\textwidth]{dir3d-z-302k_run916.png}
\caption{Director field of BPI at high shear rate: The pictures show a region 
of $L_x\times L_y \times L_Z= 32^3$ lattice sites with the flow/gradient 
direction oriented along the horizontal/vertical axis. 
The colour code gives the magnitude of the z-component of the director field. 
The data is translationally invariant in vorticity direction.}
\label{bp1-high}
\end{figure}

At high flow rates $6.25\e{-4}\lesssim\gd\lesssim7.81\e{-4}$ 
we observe a break-up of the flow-induced 
disclination network of regime BPI-2. However, 
in contrast to BPII, where the configuration is a travelling helical 
wave (viz. Fig. \ref{bp2-high}), BPI prefers a quasi 
two-dimensional, translationally invariant state, which is shown in 
Fig. \ref{bp1-high}. The snapshots visualise the affine 
transformation of simple shear flow in the gradient-flow plane.
There are distinctive regions where the director field is 
predominantly oriented along the vorticity direction and 
other regions where it is oriented rather in flow-gradient plane. 
Occasionally while moving past each other these rolls resemble 
regions of local double twist.
Interestingly the critical Ericksen number for the break-up lies in 
about the same window as that of transition between BPII-1 and BPII-2.

\subsection{Transition to cholesteric helix and flow-aligned nematic state}\label{cholflow}

At shear rates beyond the regimes BPI-3 and BPII-2 and below the transition to a 
flow-aligned nematic state we found another regime where 
both blue phases adopt the same configuration in steady shear flow, 
independent of their initial state.
The director field of this configuration is shown in Fig. \ref{cholflow}.
It consists of a simple cholesteric helix with the helical axis oriented 
along the gradient direction. The state is translational invariant in 
flow and vorticity direction and therefore also one-dimensional. 
However, in contrast to the travelling helical wave of BPII-2 the 
director field is now static and there is no tumbling motion. 
While flowing the director retains its relative orientation.
Consequently, the shear stress is lower than for
the 2D double-twist rolls of BPI-3 and the travelling helical wave
(Fig. \ref{bp2-rheo} and \ref{bp1-rheo}) as there is less dissipation.

\begin{figure}[htpb]
\includegraphics[width=0.45\textwidth]{dir3d+y-200k_run1179.png}
\caption{Flow induced cholesteric state: The orientation of the helix is 
along the gradient direction. The director field is quasi-static during 
the flow, which occurs in 'nematic layers'. For clarity the picture shows 
the director field along the y-direction for selected sites in the x-z-plane.}
\label{cholflow}
\end{figure}

%\clearpage

@@

We were able to characterise the rheology of cubic BPs and identified 
two different flow regimes for BPII and three different ones in case of BPI. 
Below Ericksen numbers $Er\simeq3$ BPII exhibits weak shear-thinning and 
obeys a power-law.  The BPII disclination network breaks up and reconnects 
in the flow, which leads to a periodically recurring dependence 
of the deviatoric shear stress. The flow-induced conformation look generally
very similar to the quiescent network at equilibrium. 
While being affinely transformed due to the shear flow
the disclination network moves in the vorticity direction. The sense of
motion is directly linked to the helicity of the underlying cholesteric phase.

At larger Ericksen numbers $4\lesssim Er\lesssim 10$ the flowing BPII network 
breaks up in a two-step scenario. First a simple cholesteric helix forms with the helical axis 
in vorticity direction. The flow is in vorticity mode, i.e. as a travelling helical wave. 
At slightly larger shear rates another cholesteric conformation emerges.
However, this time the orientation of the helix is in gradient direction.
The flow is in quasi-nematic layers with static orientation of the directior field.

Interestingly, BPI shows a flow behaviour that is very different to that of BPII.
This is a direct consequence of the topological differences in both disclination 
network. Below Ericksen numbers $Er<0.4$ the BPI network cannot flow and 
retain a regular appearance. It dissolves eventually into an amorphous network and 
features yield-stress behaviour. The reason is that shortly after the onset of the shearing 
very large stress fluctuations occur. At the lowest shear rates
the contribution of the network to the total deviatoric stress is temporarily 
actually negative during one cycle. These fluctuations destabilise the network and 
eventually trigger its dissolution into an amorphous state with 
dynamic yield stress.

At slightly larger Ericksen numbers $0.4<{\it Er}<8$ shear thinning is 
slightly stronger than for BPII, but BPI does not show simple power-law 
behaviour. 
Instead, it appears BPI kinematically bypasses the large stress fluctuations and 
flows with very complex, periodically recurring flow-induced conformations.
The offset of the network after each cycle is random in magnitude and direction.
However, despite their complex appearance the flow-induced conformations are 
topologically connected to the quiescent BPI and switching off of the 
shear flow often restores a perfect equilibrium.
It is possible to interpret these findings as an instance of 
deterministic rheochaos.

Just as in case of BPII, the BPI network breaks up at larger Ericksen numbers in
a two-step process. First at $10\lesssim\gd\lesssim 15$, it adopts a 
quasi-two-dimensional configuration that is translationally invariant in vorticity direction. 
Then at $16\lesssim Er \lesssim 22$ it forms cholestric helix along the gradient 
direction that emerges from BPII.  
Finally, at $Er\gtrsim30$ (BPI) and $Er\gtrsim10$ (BPII) the configuration
is a flow-aligned nematic state.
